A gaming peripheral brand wants to partner with Twitch streamers for a product launch. Your task: use the English-language Twitch mutual-follow network to identify the most strategically positioned streamers.


```{r}
install.packages("tidyverse")
install.packages("tidygraph")
install.packages("ggraph")
install.packages("igraph")
install.packages("tidytext")
library(tidytext)

library(tidyverse)
library(tidygraph)
library(ggraph)
library(igraph)
```


Question 1. Visualize the network and describe its structure - how many nodes, edges, and connected components?
Plot the degree distribution. Does it follow a power law?
```{r cars}
#load data and specify column types
edges <- read_csv("twitch_edges.csv", col_types = cols(from = col_double(), to = col_double()))
nodes <- read_csv("twitch_nodes.csv", col_types = cols(
  id = col_double(),
  days = col_double(),
  views = col_double(),
  new_id = col_double(),
  mature = col_logical(),
  partner = col_logical()
))
```


```{r pressure, echo=FALSE}

# Put new_id as the first column so igraph uses it as the vertex identifier
nodes_reordered <- nodes |> select(new_id, everything())

# Build graph
g <- igraph::graph_from_data_frame(edges, directed = FALSE, vertices = nodes_reordered) |>
  tidygraph::as_tbl_graph()

```

```{r}
# build a basic structure
cat("Nodes:", igraph::vcount(g), "\n")
cat("Edges:", igraph::ecount(g), "\n")
cat("Connected components:", igraph::components(g)$no, "\n")
cat("Largest component size:", max(igraph::components(g)$csize), "\n")
```

```{r}
# Filter to degree >= 10
# Have to filter because there are so many nodes to visualize
g_core <- g |>
  tidygraph::activate(nodes) |>
  mutate(degree = centrality_degree()) |>
  filter(degree >= 10)

ggraph(g_core, layout = "fr") +
  geom_edge_link(alpha = 0.1, color = "grey60") +
  geom_node_point(aes(size = degree), color = "steelblue", alpha = 0.7) +
  scale_size_continuous(range = c(0.5, 4)) +
  theme_graph() +
  labs(title = "Twitch Core Network (degree ≥ 10)")
```
Question 2. Compute degree, betweenness, and eigenvector centrality. List the top-5 streamers by each metric. Do
the three rankings agree on the same names

```{r}
# Filter to core first (speeds up computation significantly)
g_core <- g |>
  tidygraph::activate(nodes) |>
  mutate(degree = centrality_degree()) |>
  filter(degree >= 10)

# Compute all three centrality metrics
centrality_df <- g_core |>
  tidygraph::activate(nodes) |>
  mutate(
    degree      = centrality_degree(),
    betweenness = centrality_betweenness(normalized = TRUE),
    eigen       = centrality_eigen()
  ) |>
  as_tibble()

# Top 5 by degree
centrality_df |>
  select(name, degree, betweenness, eigen, partner) |>
  arrange(desc(degree)) |>
  head(5) |>
  knitr::kable(caption = "Top 5 by Degree")

# Top 5 by betweenness
centrality_df |>
  select(name, degree, betweenness, eigen, partner) |>
  arrange(desc(betweenness)) |>
  head(5) |>
  knitr::kable(caption = "Top 5 by Betweenness")

# Top 5 by eigenvector
centrality_df |>
  select(name, degree, betweenness, eigen, partner) |>
  arrange(desc(eigen)) |>
  head(5) |>
  knitr::kable(caption = "Top 5 by Eigenvector")
```
```{r}
colnames(centrality_df)
```

3. Join centrality scores with node attributes. Are official Twitch partners (is_partner) more central
than non-partners? Does views correlate more strongly with degree or eigenvector centrality? Does
views_per_day (audience efficiency) correlate differently than raw views?
```{r}
# Add views_per_day to centrality_df
centrality_df <- centrality_df |>
  mutate(views_per_day = views / days)

# --- Partner vs Non-Partner ---
centrality_df |>
  group_by(partner) |>
  summarise(
    mean_degree  = mean(degree),
    median_degree = median(degree),
    mean_eigen   = mean(eigen),
    median_eigen = median(eigen),
    n = n()
  ) |>
  knitr::kable(caption = "Partner vs Non-Partner Centrality")

# --- Correlations: views vs centrality ---
centrality_df |>
  summarise(
    views_x_degree     = cor(views, degree),
    views_x_eigen      = cor(views, eigen),
    vpd_x_degree       = cor(views_per_day, degree),
    vpd_x_eigen        = cor(views_per_day, eigen)
  ) |>
  knitr::kable(caption = "Correlations with Centrality")

# --- Visualise: views vs degree ---
ggplot(centrality_df, aes(x = degree, y = views, color = partner)) +
  geom_point(alpha = 0.4, size = 1) +
  scale_y_log10() +
  scale_x_log10() +
  labs(
    title = "Views vs Degree Centrality",
    x = "Degree (log)", y = "Views (log)",
    color = "Partner"
  ) +
  theme_minimal()

#the graph shows clear distinct clusters that confirm that partners are 2x more central and have much larger audiences
```
4. Detect communities. How many communities form? Compare the two largest on median views,
views_per_day, and days - do they represent distinct streamer profiles (e.g., established high-volume
vs. newer fast-growing)?

```{r}
install.packages("igraph")
library(igraph)

# Community detection using walktrap
g_core_undirected <- igraph::as.undirected(g_core |> igraph::as.igraph(), mode = "collapse")

g_communities <- tidygraph::as_tbl_graph(g_core_undirected) |>
  tidygraph::activate(nodes) |>
  mutate(
    community = as.factor(group_walktrap()),
    views_per_day = views / days
  )

# How many communities?
community_df <- g_communities |>
  tidygraph::activate(nodes) |>
  as_tibble()

cat("Number of communities:", n_distinct(community_df$community), "\n")

# Compare top 2 communities
community_df |>
  group_by(community) |>
  summarise(
    n              = n(),
    median_views   = median(views),
    median_vpd     = median(views_per_day),
    median_days    = median(days),
    partner_pct    = mean(partner) * 100,
    mature_pct     = mean(mature) * 100
  ) |>
  arrange(desc(n)) |>
  head(2) |>
  knitr::kable(caption = "Top 2 Communities Compared")

# Visualise communities
ggraph(g_communities, layout = "fr") +
  geom_edge_link(alpha = 0.05, color = "grey60") +
  geom_node_point(aes(color = community, size = degree), alpha = 0.7) +
  scale_size_continuous(range = c(0.5, 4)) +
  theme_graph() +
  labs(title = "Twitch Core Network — Communities")
```

5. Build a composite influencer score 
views_per_day). Using views_per_day instead of raw views corrects for channel age - a streamer with
3× the views but 10× the tenure is less efficient. Apply a brand-safety filter removing mature-content
streamers (is_mature == TRUE) before scoring. Name your top 3 targets, justify your weights, and
report whether filtering changed the ranking substantially.

```{r}
#Composite score is 40% normalised degree + 30% #eigenvector + 30% normalised


# Brand-safe filter (remove mature content streamers)
g_safe <- g_communities |>
  tidygraph::activate(nodes) |>
  filter(mature == FALSE)

# Build composite score
composite_df <- g_safe |>
  tidygraph::activate(nodes) |>
  mutate(
    degree    = centrality_degree(),
    eigen     = centrality_eigen(),
    views_per_day = views / days,
    
    # Normalise each metric to 0-1
    norm_degree = (degree - min(degree)) / (max(degree) - min(degree)),
    norm_eigen  = (eigen  - min(eigen))  / (max(eigen)  - min(eigen)),
    norm_vpd    = (views_per_day - min(views_per_day)) / (max(views_per_day) - min(views_per_day)),
    
    # Composite: 40% degree, 30% eigen, 30% views_per_day
    composite = 0.4 * norm_degree + 0.3 * norm_eigen + 0.3 * norm_vpd
  ) |>
  as_tibble()

# Top 3 targets
composite_df |>
  select(name, degree, eigen, views_per_day, views, days, partner, composite) |>
  arrange(desc(composite)) |>
  head(3) |>
  knitr::kable(caption = "Top 3 Brand-Safe Influencer Targets")

# Did filtering change the ranking? Compare with unfiltered
composite_unfiltered <- g_communities |>
  tidygraph::activate(nodes) |>
  mutate(
    degree    = centrality_degree(),
    eigen     = centrality_eigen(),
    views_per_day = views / days,
    norm_degree = (degree - min(degree)) / (max(degree) - min(degree)),
    norm_eigen  = (eigen  - min(eigen))  / (max(eigen)  - min(eigen)),
    norm_vpd    = (views_per_day - min(views_per_day)) / (max(views_per_day) - min(views_per_day)),
    composite = 0.4 * norm_degree + 0.3 * norm_eigen + 0.3 * norm_vpd
  ) |>
  as_tibble()

cat("Top 3 unfiltered:\n")
composite_unfiltered |> arrange(desc(composite)) |> slice_head(n = 3) |> pull(name) |> print()

cat("Top 3 brand-safe:\n")
composite_df |> arrange(desc(composite)) |> slice_head(n = 3) |> pull(name) |> print()

#community 4949 was filtered out by the mature filter
```
6. Geographic network. A retail chain stores adjacency data as a column listing nearby store IDs
within 10 km. In code or pseudocode: (a) convert to a tbl_graph, (b) compute degree and betweenness
centrality, (c) test whether higher-degree stores generate higher average daily sales. Interpret the
direction of the correlation.

```{r}
# (a) Convert to tbl_graph
# Assume stores dataframe has columns: store_id, avg_daily_sales, nearby_ids
# where nearby_ids is a comma-separated string of store_ids within 10km

store_edges <- stores |>
  select(Store, nearby_stores) |>
  mutate(nearby_stores = as.character(nearby_stores)) |>  # force to character first
  tidyr::separate_rows(nearby_stores, sep = ",") |>
  mutate(nearby_stores = as.integer(str_trim(nearby_stores))) |>  # then convert to integer
  rename(from = Store, to = nearby_stores) |>
  filter(!is.na(to))

g_stores <- tidygraph::tbl_graph(
  nodes = stores,
  edges = store_edges,
  directed = FALSE,
  node_key = "Store"  # match your actual ID column
)

g_stores <- g_stores |>
  tidygraph::activate(nodes) |>
  mutate(
    degree      = centrality_degree(),
    betweenness = centrality_betweenness(normalized = TRUE)
  )

store_metrics <- g_stores |>
  tidygraph::activate(nodes) |>
  as_tibble()

# Check what columns you have
colnames(store_metrics)
```

7.GraphRAG-lite. Using Data/Shared/reviews.rds, build a term co-occurrence knowledge graph
from a sample of at least 1,500 polarised reviews (stars 􀀀 2 or 􀀀 4). Tokenize with tidytext, remove
stop words, filter to terms appearing in at least 15 reviews, and draw edges between terms that co-occur
in the same review at least 5 times. Detect communities with group_walktrap(). (a) How many
communities form, and how many nodes does the largest contain? (b) List the top-6 terms by weighted
degree for the two largest communities. (c) Name each community based on its terms. (d) Which
community would you prioritise for a seeding campaign targeting dissatisfied customers, and why?

```{r}

# Load reviews
reviews <- read_csv("review.csv")

# Sample at least 1500 polarised reviews (stars <= 2 or >= 4)
reviews_polarised <- reviews |>
  filter(stars <= 2 | stars >= 4) |>
  slice_sample(n = 1500)

# Tokenize and remove stop words
tokens <- reviews_polarised |>
  select(review_id, text) |>
  tidytext::unnest_tokens(word, text) |>
  anti_join(tidytext::stop_words, by = "word") |>
  filter(!str_detect(word, "^[0-9]+$"))  # remove pure numbers

# Filter to terms appearing in at least 15 reviews
word_counts <- tokens |>
  group_by(word) |>
  summarise(n_reviews = n_distinct(review_id)) |>
  filter(n_reviews >= 15)

tokens_filtered <- tokens |>
  filter(word %in% word_counts$word)

# Build co-occurrence edges (terms that co-occur in same review >= 5 times)
co_occur <- tokens_filtered |>
  inner_join(tokens_filtered, by = "review_id") |>
  filter(word.x < word.y) |>  # avoid duplicates
  group_by(word.x, word.y) |>
  summarise(weight = n(), .groups = "drop") |>
  filter(weight >= 5)

# Build graph
g_rag <- tidygraph::tbl_graph(
  nodes = word_counts |> filter(word %in% c(co_occur$word.x, co_occur$word.y)),
  edges = co_occur
```

